Premature convergence is one of the most important drawbacks of \DE{}.
%
This paper proposes a novel variant of \DE{}, the \textit{Differential Evolution with Enhanced Diversity Maintenance} (\DEEDM{}),
which incorporates a novel replacement phase that is based on a recently proposed strategy (\RMDDC{}) that was successfully
used with memetic algorithms in the combinatorial optimization field.
%
In order to attain a faster convergence than in \RMDDC{} two modifications are proposed for the integration with \DE{}.
%
First, no concepts of the multi-objective field are applied, meaning that a more greedy strategy is devised.
%
Second, an elite population is incorporated.
%
Experimental validation shows the important benefits obtained by \DEEDM{} in long-term executions both when compared with a classic \DE{} variant and with
the state-of-the-art proposals.
%
\DEEDM{} requires setting a new parameter, which is called the initial distance factor.
%
The obtained quality highly depends on the proper specification of this parameter.
%
However, a fixed value could be used for properly dealing with benchmarks of the CEC 2016 and CEC 2017 competitions,
meaning that a quite robust proposal is obtained.

Several extensions might be explored to further improve our proposal.
%
First, with the aim of facilitating the application of \DEEDM{}, adaptive and/or self-adaptive schemes might be applied for setting the initial distance
factor, as well as for other \DE{} parameters.
%
Second, with the aim of reducing the number of evaluations required to attain high-quality solutions, some of the strategies that are used in the
field of optimization with expensive functions might be integrated into our proposal.
%
Finally, the incorporation of a local search engine mainly in the last stages of the optimization might bring additional benefits.
